---
title: "Summary of Primary Health Care Analysis"
format: html
editor: visual
execute:
  warning: false
project:
    execute-dir: file
---

```{r load helpers}
#| echo: FALSE
#| eval: TRUE
#| include: FALSE

# Load functions, objects and paths to data
  source("00_phc_quarto_setup.R")
```

## The Ask

What is PEPFAR's presence in Primary Health Care (PHC) Facilities?

-   What proportion of of PHC facilities in a country are covered by PEPFAR?

-   Where do recent HTS_TST_POS and TX_NEW come from in terms of PHC versus non-PHC facilities?

## Data Needed

To answer the question posed above, we need a series of facility lists that contain a number of features.

1.  First, to understand the universe of health facilities in a country (the denominator), we need a copy of an up-to-date master facility list (MFL). The `Org Unit` dataset (DATIM login required) available from [here](https://datim.zendesk.com/hc/en-us/articles/115002334246-DATIM-Data-Import-and-Exchange-Resources#p3 "Link to DATIM Concepts page containing Org Unit") contains all the facilities considered for the **Data Alignment Activity** (DAA). This dataset serves as the denominator for the project.

2.  Next, we need to know information about the types of facilities in a country. We recently learned an [effort](https://datim.zendesk.com/hc/en-us/articles/7881071655060-9-Site-Attributes "Link to Site Attributes article on DATIM") the DATIM support staff are conducting to gather site attributes. Through our contacts at SGAC, we were able to access a copy of the site attribute data for Botswana and Malawi. There are two limitations with these data. The site attributes only describe the type of a facility in a country. And the number of facilities with attribute information is not universal -- 88% completion in Botswana and 74% in Malawi.

3.  Finally, we need to know where PEPFAR is working in a country. This information is contained in the Monitoring, Evaluation and Reporting data (aka MER site-level MSD). We downloaded the MSDs from Panorama and retained all sites reporting on `HTS_TST_POS`, `TX_CURR`, and `TX_NEW`.

## Load and Review the Data

With our three data sets acquired, we can now load the data for each country and review the attributes.

```{r load data}
#| eval: TRUE
#| output: FALSE

# Botswana data assets
  bts_daa <- range_speedread(ss_bts_daa, col_types = "c") %>% 
    pivot_daa()
  bts_deou <- range_speedread(ss_bts_deou, col_types = "c")  
  bts_msd_site <- read_psd(bts_msd_site_path) %>% 
    munge_msd()
    
# Malawi data assets
  mwi_daa <- range_speedread(ss_mwi_daa, col_types = "c") %>% 
    pivot_daa()
  mwi_deou <- range_speedread(ss_mwi_deou, col_types = "c")
  mwi_msd_site <- read_psd(mwi_msd_site_path) %>% 
    munge_msd()
    

```

### Botswana Summary

With the data loaded, we can review the columns in each data set. Ideally, we need some form of an `orgunituid` to merge the data to the parent data set (`bts_deou` and `mwi_deou`). We can see that each dataset loaded contains some form of the `orgunituid`.

```{r review Botswana data}
#| echo: false
# Loop over the Botswana data and summarize each data frame
  map(list(bts_deou, bts_daa, bts_msd_site), ~names(.x))
```

We have set up script (`00_helper_functions`) of merge functions that help us merge the data frames together. The `tidylog` package also allows us to show how many records join on a 1-1 basis across the operations.

```{r merge Botswana}
# This is really what we are after with the ask
# Now bring in the site level indicator data
  bts_sites_mer <- join_datim_assets(df1 = bts_deou, df2 = bts_daa, df3 = bts_msd_site)
    
```

Botswana has **`r nrow(bts_deou)`** total health care facilities. Of that number, **`r nrow(bts_daa)`** contain attribute data and valid DATIM `orgunituids`. In `r glue::glue({metadata$curr_pd})`, PEPFAR reported testing and treatment indicators in **`r nrow(bts_msd_site)`** sites, of which **`r bts_sites_mer %>% filter(merge_status_two == "PEPFAR", merge_status == "merged") %>% distinct(orgunituid, merge_status) %>% count(merge_status) %>% pull(n)`** contain site attribute information -- **`r bts_sites_mer %>% filter(merge_status_two == "PEPFAR", merge_status != "merged") %>% distinct(orgunituid, merge_status) %>% count(merge_status) %>% pull(n)`** sites report PEPFAR indicators but do not have facility level attribute information. The table below provides a complete breakdown of facility types by PEPFAR coverage status.

```{r Botsana cross tab pepfar coverage}
# Fetch  country name for use in table headers
  cntry <- get_countryname(bts_sites_mer)

# PEPFAR footpring
  pepfar_footprint_gt(bts_sites_mer)
```

#### PEPFAR SHARE OF REPORTING BY FACILITY TYPE

Primary Health Centers report the greatest share of the selected testing and treatment indicators. For `TX_CURR` Primary Health Centers and Hospitals report the majority of results. It is important to keep in mind that much of USAID's efforts in Botswana occur at the community level and are not reflected in a facility-level analysis.

```{r Botswana reporting shares}
  
# Generate a summary table showing share of reporting by facility type

# Summary table of indicators by facility type for Q1
  bts_gt_df <- create_coverage_df(bts_sites_mer)

# Create a GT summary of the results
  create_phc_gt(bts_gt_df)

```

### Malawi Summary

We use the same process to load and merge the Malawi data. First, we'll join the three data frames together.

```{r merge Malawi}
# Merge together the three Malawi data assets 
  mwi_sites_mer <- join_datim_assets(mwi_deou, mwi_daa, mwi_msd_site)
```

As summarized in the merge results above, Malawi has `r nrow(mwi_deou)` total health care facilities. Of this number, **`r nrow(mwi_daa)`** contain facility attribute information. In `r glue::glue({metadata$curr_pd})`, PEPFAR reported testing and treatment indicators in **`r nrow(mwi_msd_site)`** sites, of which **`r mwi_sites_mer %>% filter(merge_status_two == "PEPFAR", merge_status == "merged") %>% distinct(orgunituid, merge_status) %>% count(merge_status) %>% pull(n)`** contain site attribute information -- **`r mwi_sites_mer %>% filter(merge_status_two == "PEPFAR", merge_status != "merged") %>% distinct(orgunituid, merge_status) %>% count(merge_status) %>% pull(n)`** sites report PEPFAR indicators but do not have facility level attribute information. The table below provides a complete breakdown of facility types by PEPFAR coverage status.

```{r Malawi cross tab pepfar coverage}
# Generate a summary table showing share of reporting by facility type

# Fetch  country name for use in table headers
  cntry <- get_countryname(mwi_sites_mer)

# PEPFAR footpring
  pepfar_footprint_gt(mwi_sites_mer)
```

#### PEPFAR SHARE OF REPORTING BY FACILITY TYPE

Similar to Botswana, Primary Health Centers report the greatest share of the selected testing and treatment indicators. Across all selected indicators Primary Health Centers and Hospitals report the majority of results.

```{r Malawi reporting shares}
  
# Generate a summary table showing share of reporting by facility type

# Summary table of indicators by facility type for Q1
  mwi_gt_df <- create_coverage_df(mwi_sites_mer)

# Create a GT summary of the results
  create_phc_gt(mwi_gt_df)

```

### Conclusion

According to analysis of six data sets in two operating units, we learned that Primary Health Centers and Hospitals drive the majority of MER reporting results across select testing and treatment indicators. In Botswana, only **`r bts_sites_mer %>% filter(merge_status_two == "PEPFAR", merge_status == "merged") %>% distinct(orgunituid, merge_status) %>% count(merge_status) %>% pull(n)`** facilities report on MER indicators and have facility attribute information. In Malawi, this number is much higher at **`r mwi_sites_mer %>% filter(merge_status_two == "PEPFAR", merge_status == "merged") %>% distinct(orgunituid, merge_status) %>% count(merge_status) %>% pull(n)`**.
