---
title: "Summary of Primary Health Care Analysis"
format: html
editor: visual
project:
    execute-dir: file
---

```{r}
getwd()
```

## The Ask

What is PEPFAR's presence in Primary Health Care (PHC) Facilities?

-   What proportion of of PHC facilities in a country are covered by PEPFAR?

-   Where do recent HTS_TST_POS and TX_NEW come from in terms of PHC versus non-PHC facilities?

## Data Needed & Definitions

To answer the question posed above, we need a series of facility lists that contain a number of features.

1.  First, to understand the universe of health facilities in a country (the denominator), we need a copy of an up-to-date master facility list (MFL). The `Org Unit` dataset (DATIM login required) available from here contains all the facilities

2.  Second, we need to know in which facilities PEPFAR is working. We can use the site level MER Structured datasets (or DATIM) available from Panorama to track PEPFAR's facility-level footprint in addition to distinguishing what indicators are reported at each place. The site level MSDs do not contain a column that identify facility attributes.

3.  Third, we need a common crosswalk that enables one to merge the facilities from the MFL to those contained in DATIM. Within DATIM, facilities have an `orgunituid`, `facility name`, and `facilityuid`. Ideally, the MFL would contain a common field that would enable the facility level data to be easily joined between data sets. **No MFL that we have reviewed contain a common field on which data can be reliably joined.**

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
